### 2011-10-07 tex/abstract.tex
initial commit

Explicit-state-duration hidden Markov models (EDHMM) are HMMs that have latent states consisting of both discrete state-indicator and discrete state-duration random variables.  In contrast to the implicit geometric state duration distribution possessed by the standard HMM, EDHMMs allow the direct parameterisation and estimation of per-state duration distributions. As most duration distributions have support on all positive integers, truncation or other approximations are usually required to perform EDHMM inference.  In this letter we borrow from the inference techniques developed for unbounded state-cardinality (nonparametric) variants of the HMM and use them to develop a tuning-parameter free inference approach for EDHMMs.  We illustrate the performance gains of this approach using synthetic data.