### 2011-10-07 tex/results.tex

\begin{figure}
    \subfloat[][]{
        \includegraphics[width=0.5\textwidth]{experiment_2_X.pdf}
        \label{fig:exp_1_state}
    } \\
    \subfloat[][]{
        \includegraphics[width=0.5\textwidth]{experiment_2_Y.pdf}
        \label{fig:exp_1_data}
    }
    \caption{Example a) state and b) observation sequence generated by the explicit duration HMM. Here $K$ = 3; $p(y_t|x_t=j) = \mathrm{N}(\mu_j, 1)$ with $\mu_1 = -3$, $\mu_2 = 0$, and $\mu_3 = 3$; and $p(d_t|x_t=j) = \mathrm{Poisson}(\lambda_j)$ with $\lambda_1 = 5$, $\lambda_2 = 15$, and $\lambda_3 = 20$.}
    \label{fig:experiment1_data}
\end{figure}

\begin{figure}
    \subfloat[][]{
        \includegraphics[width=0.25\textwidth]{posterior_means.pdf}
        \label{fig:posterior_means}
    } 
    \subfloat[][]{
        \includegraphics[width=0.25\textwidth]{posterior_rates.pdf}
        \label{fig:posterior_rates}
    }
    \caption{Samples from the posterior distributions of a) the observation distribution means and b) the duration distribution rate parameters.}
    \label{fig:experiment1_results}
\end{figure}
\subsection{Synthetic Data}

The first experiment uses the 500 data points shown in Figure \ref{fig:experiment1_data} generated from an EDHMM with three states. To generate the data the duration distributions were chosen to be Poisson with rates $\lambda_1 = 5$, $\lambda_2 = 15$, $\lambda_3 = 20$; each observation distribution was Gaussian with means of $\mu_1 = -3$, $\mu_2 = 0$, and $\mu_3 = 3$, each with a variance of 1. The transition distributions $A$ were set to
\begin{equation*}
\begin{bmatrix}
    0 & 0.3 & 0.7 \\ 0.6 & 0 & 0.4 \\ 0.3 & 0.7 & 0
\end{bmatrix}.  
\end{equation*}

For inference, broad, uninformative priors were chosen for the parameters of the duration and observation distributions. The parameters of the observation distribution for all states were given a normal-inverse-Wishart (N-IW) prior with parameters $\nu_0 = 2$, $\Lambda_0 = 1$, $\kappa=0.1$ and $\mu_0 = 0$. The rate parameters for all states were given $\mathrm{Gamma}(1, 10^{5})$ priors.

One thousand samples were collected from the EDHMM beam sampler after a burn-in of 500 samples. The learned posterior distribution of the state duration parameters and means of the observation distributions are shown in Figure \ref{fig:experiment1_results}.  The EDHMM achieves high accuracy in the estimated posterior distribution of the observation means, despite the overlap in observation distributions. The rate parameter distributions are reasonably estimated given the small number of observed segments.

A second experiment was performed using similar synthetic data to demonstrate the ability of the EDHMM to distinguish between states having differing duration distributions but the same observation distribution. Here the same model was used as above except data 
%was 
were
generated using $\mu_1 = 0$, $\mu_2 = 0$, and $\mu_3 = 3$. The same priors were used as above, and again the sampler was allowed to run for 1000 samples with a burn-in of 500 samples.

Figure~\ref{fig:experiment2_results} shows that the sampler clearly separates the high state associated with $\mu_3$ from the other two states and clearly 
%identifies the fact that there are 
reveals the presence of
two  low states with differing duration distributions.   That it is possible for the EDHMM to recover two states with differing duration distributions requires a third state.  Figure~\ref{fig:exp_2_state} shows posterior samples that indicate that the model is mixing over ambiguities about states $0$ and $1$ as it should. %As shown in Figure \ref{fig:experiment2_results}, when the state transitions to and from the long-high state to the short-low state (e.g. $t\approx 100$ and $t\approx 180$) the transitions are captured well. When the system transitions between the two low states the segmentation is less accurate, due to the inherent lack of identifiability during these periods. Instead, the space of valid state sequences is explored, each sequence obeying the duration distribution.

%Despite the loss of identifiability in the state sequence, the sampler is still able to capture reasonable estimates for the parameters of the observation and duration distributions.

\begin{figure}
    \subfloat[][]{
        \includegraphics[width=0.5\textwidth]{experiment_3_Y.pdf}
        \label{fig:exp_2_data}
    } \\
    \subfloat[][]{
        \includegraphics[width=0.5\textwidth]{experiment_3_X.pdf}
        \label{fig:exp_2_state}
    } \\
    \subfloat[][]{
        \includegraphics[width=0.25\textwidth]{posterior_means_exp3.pdf}
        \label{fig:posterior_means_3}
    } 
    \subfloat[][]{
        \includegraphics[width=0.25\textwidth]{posterior_rates_exp3.pdf}
        \label{fig:posterior_rates_3}
    }
    \caption{Results of the beam sampler applied to a system that generates states with identical observation distribution but differing durations. The observations are shown in a), and the true states are shown in b) overlaid with 20 randomly selected state traces produced by the sampler after burn-in. Samples from the posterior observation distribution mean are shown in c), and samples from the posterior duration distribution rates are shown in d).}
    \label{fig:experiment2_results}
\end{figure}

Finally, Figure \ref{fig:allowed} shows that the number of allowed transitions decays with the number of iterations, significantly speeding up the runtime of the later samples after the burn-in period.

\begin{figure}
    \includegraphics[width=0.5\textwidth]{number_transitions_visited.pdf}

\caption{Mean number of transitions visited over the first 1000 iterations of the beam sampler.}
\label{fig:allowed}
\end{figure}