### 2011-11-03 tex/edhmm.tex
final update

\documentclass[journal]{../inc/oldtran/IEEEtran}
\usepackage{amsmath,amssymb,graphicx,subfig}
\usepackage[usenames,dvipsnames]{color}
\usepackage{algorithm,algorithmic}
\newcommand{\E}{\textrm{E}}

\begin{document}

\title{Inference in Hidden Markov Models with Explicit State Duration Distributions}
\author{Michael Dewar, Chris Wiggins, Frank Wood}%

%\thanks{Michael Dewar is with bitly, New York City}%
%\thanks{Chris Wiggins is with the Department of Applied Physics and Applied Mathematics, Columbia University}%
%\thanks{Frank Wood is the the Department of Statistics, Columbia Univeristy}%

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}

\input{introduction}

    \begin{figure}[t]
        \centering
        \subfloat[][]{
            \includegraphics[width=0.23\textwidth]{../pic/EDHMM_graphical_model.pdf}
            \label{fig:graphical model}
        }
        \subfloat[][]{
            \includegraphics[width=0.23\textwidth]{../pic/EDHMM_aux_graphical_model.pdf}
         \label{fig:aux graphical model}
        }
        \caption{a) The Explicit Duration Hidden Markov Model. The time left in the current state $x_t$ is denoted $d_t$. The observation at each point in time is denoted $y_t$. b) The EDHMM with the additional auxiliary variable $u_t$ used in the beam sampler.}
    \label{fig:graphs}    
    \end{figure}
\section{Explicit Duration Hidden Markov Model}
\label{sec:Model}

\input{model}

\input{algorithm}

\section{EDHMM Inference}

\label{sec:inference}

\input{sampling}

\subsection{Related Work}

The need to accommodate explicit state duration distributions in HMMs has long been recognised. Rabiner \cite{Rabiner89} details the basic approach which expands the state space to include dwell time before applying a slightly modified Baum-Welch algorithm. This approach specifies a maximum state duration, limiting practical application 
to cases with short sequences and dwell times.
This approach, generalised under the name ``segmental hidden Markov models'', includes more general transitions than those Rabiner considered, allowing the next state and duration to be conditioned on the previous state and duration \cite{Gales93}. Efficient approximate inference procedures were developed in the context of speech recognition \cite{Ostendorf96} and evolved into symmetric approaches suitable for practical implementation \cite{Yu2006}.

Recently, a ``sticky'' variant of the hierarchical Dirichlet process HMM (HDP-HMM) has been developed \cite{Fox2008}.  The HDP-HMM has countable state-cardinality \cite{Teh06} allowing estimation of the number of states in the HMM; the sticky aspect addresses long dwell times by introducing a parameter in the prior that favours self-transition.

\section{Experiments}

\label{sec:experiments}

\input{results}

\section{Discussion}

\label{sec:dicussion}

\input{discussion}

\bibliographystyle{plain}

\bibliography{../inc/library}

\end{document}